Comprehensive Quality Awareness Automated Semantic Web Service Composition

### 2017-04-01 paper.tex
update

The \emph{QoS} of the service composition can be obtained by aggregating the QoS values of the participating services. For a service composition with services $ S_1, S_2, ... S_n$ we obtain the reliabilitiy $R=\prod\limits^n_{k=1}r_{S_k}$, the availability $A=\prod\limits^n_{k=1}a_{S_k}$, the cost $C=\sum\limits^n_{k=1}c_{S_k}$, and the response time $T$ is the time of most time-consumption path in the composition, i.e., $$T = MAX \{\sum\limits^{\ell_j}_{k=1}t_{S_k} | \text{ $j\in\{1,\ldots,m\}$ and $P_j$ is a path of length $\ell_j$}\}.$$

@@

\begin{algorithm}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{createWeightedDAG}{createWeightedDAG}
 \SetKwProg{Procedure}{Procedure}{}{}
 %\LinesNumbered
 \SetNlSty{}{}{:}
 % \Procedure{}{
 \Input{ $I_T$, $O_T$, $queue$, $cache$}
 \Output{WG}
 $WG = (V, E)$\;
 $V \leftarrow$ \{$Start$, $End$ \} \;
 $OutputSet \leftarrow$ \{$I_{T}$\}\;
  \While { $O_{T}$ do not satisfied by $OutputSet$}{
     \ForEach{$S$ in $queue$}{
      \uIf{all $I_{S}$ satisfied by $OutputSet$}{    
        $e \leftarrow$ calcuate $type_e$, $sim_e$\;
        $E$ add $e$\;
        $V$ add $S$\;
        $OutputSet$ add \{$O_{S}$\}\;    
        $queue$.remove $S$\;
       }	
     }
  }
 remove $dangling nodes$\; 
 remove $dangling edges$\;
 \KwRet $WG$\;
 %}
 \caption{Create a weighted DAG from a  sorted queue.}
\label{graph_building}
\end{algorithm}

@@

\section{Experiment Study}\label{experiment_design}
In this section, we employ a quantitative evaluation approach with a benchmark dataset used in \cite{ma2015hybrid,da2016genetic}, an augmented version of Web Service Challenge 2009(wsc09) including QoS attributes. Two objectives of this evaluation are to: $(1)$ measure the effectiveness of our PSO-based approach, see comparision test in \ref{comparisonTestWithGP}.
%To do this, we compare our PSO-based method with one existing GP-based approach \cite{ma2015hybrid}. 
$(2)$ measure the effectiveness of our proposed comprehensive quality to achieve a desirable balance on semantic matchmaking quality and QoS, see comparision test in \ref{comparisonTest}.
%To do this, we compare a widely used QoS model in QoS-aware web service composition \cite{ma2015hybrid,da2016particle,da2015graphevol} with our comprehensive quality model using our PSO-based approach.

@@

The parameters for the PSO are chosen from the settings from \cite{shi2001particle}, In particular, PSO population size is 30 with 100 generations. We run 30 times independently for each dataset. We configure the weights of fitness function to properly balance semantic matchmaking quality and QoS. Therefore, $w_{1}$ and $w_{2}$ are set equally to 0.25, and $w_{3}$, $w_{4}$, $w_{5}$, $w_{6}$ are all set to 0.125. The $p$ of $type_e$ is set to 0.75 ($plugin match$) according to \cite{lecue2009optimizing}. In general, weight settings and parameter $p$ are decided by users' preferences.
\vspace{-0.5cm}
\subsection{Comparison Test for GP-based approach and PSO-based approach}\label{comparisonTestWithGP}
To evaluate the effectiveness of our proposed PSO-based approach, we compare one recent GP-based approach \cite{ma2015hybrid} with our PSO-based method using our proposed comprehensive quality model. We extend the GP-based approach by measuring the semantic matchmaking quality between parent nodes and children nodes. To make a fair comparison, we use the same number of evaluations (3000 times) for these two approach. We set the parameters' settings of that GP-based approach as 30 individuals and 100 generations, it is considered to be proper settings refering to \cite{da2015gp}.

The first column of Table \ref{meanFitness} shows five tasks from WSC09. The second and third column of Table \ref{meanFitness} show the orginal service repository size before the greedy search and the shrunk service repository size after the greedy search respectively regarding the five tasks. This greedy search helps reducing the original repository size significantly, which contributes to a reduced searching space. The fourth and fifth column of Table \ref{meanFitness} show the mean fitness values of 30 independent runs accomplished by two methods. We employ independent-samples T tests to test the significant differences in mean fitness value. The results show that the PSO-based approach outperformes the existing GP-based approach in most cases except task 3 (all the p-values are consistently smaller than 0.01). In task 5, the PSO-based approach performs significantly better than the GP-based approach in finding optimal solutions. It may be that the GP-based approach is stuck in local optima due to the very large search space in Task 5. On the other hand, the decoding process used by the PSO-based approach allows for small changes that more effectively prevent this from happening.
\vspace{-0.8cm}
\begin{table}[]
\centering
\caption{Mean fitness values for comparing GP-based approach}
\label{meanFitness}
\begin{tabular}{c|c|c|l|l}
\hline
\multicolumn{1}{c|}{WSC09} &Original $\mathcal{SR}$  &Shrunken $\mathcal{SR}$   &PSO-based approach & GP-based approach  \\ \hline
Task 1                     &572            &80    &0.5592 $\pm$ 0.0128  $\uparrow$  &0.5207 $\pm$ 0.0208           \\ \hline
Task 2                     &4129           &140   &0.4701 $\pm$ 0.0011  $\uparrow$  &0.4597 $\pm$ 0.0029          \\ \hline
Task 3                     &8138           &153   &0.5504 $\pm$ 0.0128              &0.5679 $\pm$ 0.0234 $\uparrow$   \\ \hline
Task 4                     &8301           &330   &0.4690 $\pm$ 0.0017  $\uparrow$  &0.4317 $\pm$ 0.0097            \\ \hline
Task 5                     &15211          &237   &0.4694 $\pm$ 0.0008  $\uparrow$  &0.2452 $\pm$ 0.0369            \\ \hline
\end{tabular}
\end{table}
\vspace{-0.8cm}

@@

We observe an interesting pattern from Table \ref{decisionTable}. The mean values of $QoS$ using QoS evaluation model are significantly higher than those using comprehensive quality evaluation model for Tasks 2, 3, 4 and 5. However, the mean value of $SM$ using the comprehensive quality evaluation model are significantly higher than those using the QoS evaluation model, while a slight trade-off in $QoS$ are observed in all tasks.
\vspace{-0.8cm}
\begin{table}[]
\footnotesize
\centering
\caption{Mean values of $SM$ and $QoS$ for QoS evaluation model and comprehensive quality evaluation model using PSO-based approach}
\label{decisionTable}
\begin{tabular}{c|l|l|l}
\hline
\multicolumn{2}{c|}{WSC09}              & \shortstack{QoS \\ Evaluation Model}         &\shortstack{Comprehensive Quality \\ Evaluation Model} \\ \hline
\multirow{2}{*}{Task1}  &$SM$   &0.5373 $\pm$ 0.0267               &0.5580 $\pm$ 0.0094 $\uparrow$ \\ \cline{2-4}
                        &$QoS$  &0.5574 $\pm$ 0.0156               &0.5604 $\pm$ 0.0164                          \\ \hline
\multirow{2}{*}{Task2}  &$SM$   &0.4549 $\pm$ 0.0033               &0.4630 $\pm$ 0.0042 $\uparrow$ \\ \cline{2-4} 
                        &$QoS$  &0.4800 $\pm$ 0.0012 $\uparrow$    &0.4772 $\pm$ 0.0025 \\ \hline
\multirow{2}{*}{Task3}  &$SM$   &0.5538 $\pm$ 0.0082               &0.6093 $\pm$ 0.0054 $\uparrow$   \\ \cline{2-4} 
                        &$QoS$  &0.4940 $\pm$ 0.0013 $\uparrow$    &0.4913 $\pm$ 0.0009            \\ \hline
\multirow{3}{*}{Task4}  &$SM$   &0.4398 $\pm$ 0.0037               &0.4604 $\pm$ 0.0000 $\uparrow$ \\ \cline{2-4} 
                        &$QoS$  &0.4845 $\pm$ 0.0010 $\uparrow$    &0.4734 $\pm$ 0.0044  \\ \hline
\multirow{3}{*}{Task5}  &$SM$   &0.4580 $\pm$ 0.0065               &0.4639 $\pm$ 0.0013 $\uparrow$           \\ \cline{2-4} 
                        &$QoS$  &0.4764 $\pm$ 0.0005 $\uparrow$    &0.4750 $\pm$ 0.0007  \\ \hline                                                   
\end{tabular}
\end{table}
\vspace{-0.8cm}
\subsection{Further Discussion}\label{discuss1}
To analyze the effectiveness of achieving a good comprehensive quality at the expense of slightly reduced QoS, we demonstrate two best solutions produced using task 3 as an example. Fig. \ref{comparisontest} $(1)$ and $(2)$ show two $WGs$ obtained by employing the QoS evaluation model and the comprehensive quality evaluation model respectively. Both weighted DAGs have exactly the same service workflow structure, but some service vertices and edges denoted in red are different. To better understand these differences, we list the overall semantic matchmaking quality $SM$,  overall $QoS$ and semantic matchmaking quality associated to each edge ($sm_{e_1}$ to $sm_{e_4}$) in Fig. \ref{comparisontest} $(3)$, where $\Delta Q$ reveals the gain (positive $\Delta Q$) or a loss (negative $\Delta Q$) of the listed qualities for our comprehensive quality evalutaion model. Therefore, an overall gain 0.1433 is calculated from a sum of a $SM$ gain (0.1467) and $QoS$ loss (-0.0034) using comprehensive quality evaluation model. To understand the improvement of semantic matchmaking quality from these numbers, we pick up $e_4$ that is associated with the smallest $\Delta Q$. The $e_4$ of Fig. \ref{comparisontest} $(1)$ and Fig. \ref{comparisontest} $(2)$ has two different source service vertices, $Ser1640238160$ and $Ser947554374$, and the same $end$ vertices. $Ser1640238160$ and $Ser947554374$ are services with concept-related output parameters $Inst795998200$ and $Inst582785907$ corresponds to two concepts $Con103314376$ and $Con2037585750$ respectively in the ontology shown in Fig. \ref{comparisontest} $(4)$. In addtion, $Inst658772240$ is a required parameter of the $end$ vertice, which is related to concept $Con2113572083$. Obviouly,  $Inst795998200$ is closer to user's required output $Inst658772240$ compared to $Inst582785907$.
\vspace{-0.6cm}

\begin{figure}[h]
\centering{
\fbox{
\includegraphics[scale=.29]{comparisontest.pdf}}}
 \caption{An example of comparision to optimal solutions using Task 3 for QoS evaluation model and comprehensive quality evluation model.}
 \label{comparisontest}
\end{figure}
\vspace{-1.0cm}

@@

In this work, we propose a PSO-based approach to comprehensive quality-aware semantic web service composition, which produces serivce composition solutions with a desirable balance on both semantic matchmaking quality and QoS. Future works can investigate multi-objective EC techniques to produce a set of composition solutions for the situations when the quality preference is not known.

>>>>

The \emph{QoS} of the service composition can be obtained by aggregating the QoS values of the participating services. For a service composition with services $ S_1, S_2, ... S_n$ we obtain the reliability $R=\prod\limits^n_{k=1}r_{S_k}$, the availability $A=\prod\limits^n_{k=1}a_{S_k}$, the cost $C=\sum\limits^n_{k=1}c_{S_k}$, and the response time $T$ is the time of most time-consumption path in the composition, i.e., $$T = MAX \{\sum\limits^{\ell_j}_{k=1}t_{S_k} | \text{ $j\in\{1,\ldots,m\}$ and $P_j$ is a path of length $\ell_j$}\}.$$

@@

\begin{algorithm}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{createWeightedDAG}{createWeightedDAG}
 \SetKwProg{Procedure}{Procedure}{}{}
 %\LinesNumbered
 \SetNlSty{}{}{:}
 % \Procedure{}{
 \Input{ $I_T$, $O_T$, $queue$, $cache$}
 \Output{WG}
 $WG = (V, E)$\;
 $V \leftarrow$ \{$Start$, $End$ \} \;
 $OutputSet \leftarrow$ \{$I_{T}$\}\;
  \While { $O_{T}$ do not satisfied by $OutputSet$}{
     \ForEach{$S$ in $queue$}{
      \uIf{all $I_{S}$ satisfied by $OutputSet$}{    
        $e \leftarrow$ calculate $type_e$, $sim_e$\;
        $E$ add $e$\;
        $V$ add $S$\;
        $OutputSet$ add \{$O_{S}$\}\;    
        $queue$.remove $S$\;
       }	
     }
  }
 remove $dangling nodes$\; 
 remove $dangling edges$\;
 \KwRet $WG$\;
 %}
 \caption{Create a weighted DAG from a  sorted queue.}
\label{graph_building}
\end{algorithm}

@@

\section{Experiment Study}\label{experiment_design}
In this section, we employ a quantitative evaluation approach with a benchmark dataset used in \cite{ma2015hybrid,da2016genetic}, an augmented version of Web Service Challenge 2009(wsc09) including QoS attributes. Two objectives of this evaluation are to: $(1)$ measure the effectiveness of our PSO-based approach, see comparison test in \ref{comparisonTestWithGP}.
%To do this, we compare our PSO-based method with one existing GP-based approach \cite{ma2015hybrid}. 
$(2)$ measure the effectiveness of our proposed comprehensive quality to achieve a desirable balance on semantic matchmaking quality and QoS, see comparison test in \ref{comparisonTest}.
%To do this, we compare a widely used QoS model in QoS-aware web service composition \cite{ma2015hybrid,da2016particle,da2015graphevol} with our comprehensive quality model using our PSO-based approach.

@@

The parameters for the PSO are chosen from the settings from \cite{shi2001particle}, In particular, PSO population size is 30 with 100 generations. We run 30 times independently for each dataset. We configure the weights of fitness function to properly balance semantic matchmaking quality and QoS. Therefore, $w_{1}$ and $w_{2}$ are set equally to 0.25, and $w_{3}$, $w_{4}$, $w_{5}$, $w_{6}$ are all set to 0.125. The $p$ of $type_e$ is set to 0.75 ($plugin match$) according to \cite{lecue2009optimizing}. In general, weight settings and parameter $p$ are decided by users' preferences.
\vspace{-0.5cm}
\subsection{Comparison Test for GP-based approach and PSO-based approach}\label{comparisonTestWithGP}
To evaluate the effectiveness of our proposed PSO-based approach, we compare one recent GP-based approach \cite{ma2015hybrid} with our PSO-based method using our proposed comprehensive quality model. We extend the GP-based approach by measuring the semantic matchmaking quality between parent nodes and children nodes. To make a fair comparison, we use the same number of evaluations (3000 times) for these two approach. We set the parameters' settings of that GP-based approach as 30 individuals and 100 generations, it is considered to be proper settings referring to \cite{da2015gp}.

The first column of Table \ref{meanFitness} shows five tasks from WSC09. The second and third column of Table \ref{meanFitness} show the original service repository size before the greedy search and the shrunk service repository size after the greedy search respectively regarding the five tasks. This greedy search helps reducing the original repository size significantly, which contributes to a reduced searching space. The fourth and fifth column of Table \ref{meanFitness} show the mean fitness values of 30 independent runs accomplished by two methods. We employ independent-samples T tests to test the significant differences in mean fitness value. The results show that the PSO-based approach outperforms the existing GP-based approach in most cases except task 3 (all the p-values are consistently smaller than 0.01). In task 5, the PSO-based approach performs significantly better than the GP-based approach in finding optimal solutions. It may be that the GP-based approach is stuck in local optima due to the very large search space in Task 5. On the other hand, the decoding process used by the PSO-based approach allows for small changes that more effectively prevent this from happening.
\vspace{-0.8cm}
\begin{table}[]
\centering
\caption{Mean fitness values for comparing GP-based approach}
\label{meanFitness}
\begin{tabular}{c|c|c|l|l}
\hline
\multicolumn{1}{c|}{WSC09} &Original $\mathcal{SR}$  &Shrunken $\mathcal{SR}$   &PSO-based approach & GP-based approach  \\ \hline
Task 1                     &572            &80    &0.5592 $\pm$ 0.0128  $\uparrow$  &0.5207 $\pm$ 0.0208           \\ \hline
Task 2                     &4129           &140   &0.4701 $\pm$ 0.0011  $\uparrow$  &0.4597 $\pm$ 0.0029          \\ \hline
Task 3                     &8138           &153   &0.5504 $\pm$ 0.0128              &0.5679 $\pm$ 0.0234 $\uparrow$   \\ \hline
Task 4                     &8301           &330   &0.4690 $\pm$ 0.0017  $\uparrow$  &0.4317 $\pm$ 0.0097            \\ \hline
Task 5                     &15211          &237   &0.4694 $\pm$ 0.0008  $\uparrow$  &0.2452 $\pm$ 0.0369            \\ \hline
\end{tabular}
\end{table}
\vspace{-0.8cm}

@@

We observe an interesting pattern from Table \ref{decisionTable}. The mean values of $QoS$ using QoS evaluation model are significantly higher than those using comprehensive quality evaluation model for Tasks 2, 3, 4 and 5. However, the mean value of $SM$ using the comprehensive quality evaluation model are significantly higher than those using the QoS evaluation model, while a slight trade-off in $QoS$ are observed in all tasks.
\vspace{-0.8cm}
\begin{table}[]
\footnotesize
\centering
\caption{Mean values of $SM$ and $QoS$ for QoS evaluation model and comprehensive quality evaluation model using PSO-based approach}
\label{decisionTable}
\begin{tabular}{c|l|l|l}
\hline
\multicolumn{2}{c|}{WSC09}              & \shortstack{QoS \\ Evaluation Model}         &\shortstack{Comprehensive Quality \\ Evaluation Model} \\ \hline
\multirow{2}{*}{Task1}  &$SM$   &0.5373 $\pm$ 0.0267               &0.5580 $\pm$ 0.0094 $\uparrow$ \\ \cline{2-4}
                        &$QoS$  &0.5574 $\pm$ 0.0156               &0.5604 $\pm$ 0.0164                          \\ \hline
\multirow{2}{*}{Task2}  &$SM$   &0.4549 $\pm$ 0.0033               &0.4630 $\pm$ 0.0042 $\uparrow$ \\ \cline{2-4} 
                        &$QoS$  &0.4800 $\pm$ 0.0012 $\uparrow$    &0.4772 $\pm$ 0.0025 \\ \hline
\multirow{2}{*}{Task3}  &$SM$   &0.5538 $\pm$ 0.0082               &0.6093 $\pm$ 0.0054 $\uparrow$   \\ \cline{2-4} 
                        &$QoS$  &0.4940 $\pm$ 0.0013 $\uparrow$    &0.4913 $\pm$ 0.0009            \\ \hline
\multirow{3}{*}{Task4}  &$SM$   &0.4398 $\pm$ 0.0037               &0.4604 $\pm$ 0.0000 $\uparrow$ \\ \cline{2-4} 
                        &$QoS$  &0.4845 $\pm$ 0.0010 $\uparrow$    &0.4734 $\pm$ 0.0044  \\ \hline
\multirow{3}{*}{Task5}  &$SM$   &0.4580 $\pm$ 0.0065               &0.4639 $\pm$ 0.0013 $\uparrow$           \\ \cline{2-4} 
                        &$QoS$  &0.4764 $\pm$ 0.0005 $\uparrow$    &0.4750 $\pm$ 0.0007  \\ \hline                                                   
\end{tabular}
\end{table}
\vspace{-0.8cm}
\subsection{Further Discussion}\label{discuss1}
To analyse the effectiveness of achieving a good comprehensive quality at the expense of slightly reduced QoS, we demonstrate two best solutions produced using task 3 as an example. Fig. \ref{comparisontest} $(1)$ and $(2)$ show two $WGs$ obtained by employing the QoS evaluation model and the comprehensive quality evaluation model respectively. Both weighted DAGs have exactly the same service workflow structure, but some service vertices and edges denoted in red are different. To better understand these differences, we list the overall semantic matchmaking quality $SM$,  overall $QoS$ and semantic matchmaking quality associated to each edge ($sm_{e_1}$ to $sm_{e_4}$) in Fig. \ref{comparisontest} $(3)$, where $\Delta Q$ reveals the gain (positive $\Delta Q$) or a loss (negative $\Delta Q$) of the listed qualities for our comprehensive quality evaluation model. Therefore, an overall gain 0.1433 is calculated from a sum of a $SM$ gain (0.1467) and $QoS$ loss (-0.0034) using comprehensive quality evaluation model. To understand the improvement of semantic matchmaking quality from these numbers, we pick up $e_4$ that is associated with the smallest $\Delta Q$. The $e_4$ of Fig. \ref{comparisontest} $(1)$ and Fig. \ref{comparisontest} $(2)$ has two different source service vertices, $Ser1640238160$ and $Ser947554374$, and the same $end$ vertices. $Ser1640238160$ and $Ser947554374$ are services with concept-related output parameters $Inst795998200$ and $Inst582785907$ corresponds to two concepts $Con103314376$ and $Con2037585750$ respectively in the ontology shown in Fig. \ref{comparisontest} $(4)$. In addition, $Inst658772240$ is a required parameter of the $end$ service, and related to concept $Con2113572083$. Obviously,  $Inst795998200$ is closer to user's required output $Inst658772240$ compared to $Inst582785907$.
\vspace{-0.6cm}

\begin{figure}[h]
\centering{
\fbox{
\includegraphics[scale=.29]{comparisontest.pdf}}}
 \caption{An example of comparison to optimal solutions using Task 3 for QoS evaluation model and comprehensive quality evluation model.}
 \label{comparisontest}
\end{figure}
\vspace{-1.0cm}

@@

In this work, we propose a PSO-based approach to comprehensive quality-aware semantic web service composition, which produces service composition solutions with a desirable balance on both semantic matchmaking quality and QoS. Future works can investigate multi-objective EC techniques to produce a set of composition solutions for the situations when the quality preference is not known.
